### 2015-03-05 Paper/scratch.tex

    \draw [->,decorate,decoration={snake, post length=1mm}] (-3.0,0.0)
    -- (-0.1,0.0) node[midway,above] {$\epsilon=h\nu$} ; 
    \draw [->,decorate,decoration={snake, post length=1mm}] (0.1,0.1)
    -- (2.5,1.7) node[midway,sloped,above] {$\epsilon_1=h\nu_1$};
    \draw (1.0,0.0) arc (0:37:1cm);
    \draw (1.2,0.35) node {$\theta$};
    \draw (1.3,-0.7) node {$\mathbf{p},E$};
    \fill (0,0) circle (2pt);
    \draw [->] (0,0) -- (1.5,-1.5);
    \draw (1.75,-1.75) node {$\vec{P}_{ef}$};
    \draw (-3.35,0.0) node {$\vec{P}_{\gamma i}$};
    \draw (2.8,1.85) node {$\vec{P}_{\gamma f}$};

, and imgine further than
the angle between the jet direction and the observer line of sight is
$i$.

What are the advantages of this method of flagging LISA sources?  Can
you go down to binaries that cannot be resolved with VLBI? Can you go
down in mass or mass ratio or something?

  = 2.42\, \mathrm{degrees}.