### 2015-07-06 Paper/paper.tex
Made more changes suggested by Avi.

\documentclass[iop]{emulateapj}
\bibliographystyle{apj}
\usepackage{amsmath} 
\usepackage{aas_macros,amssymb}
\usepackage{natbib,graphicx,layouts}
\usepackage{color,hyperref}
\usepackage{bm}
\definecolor{linkcolor}{rgb}{0,0,0.5}
\definecolor{notecolor}{rgb}{0.8,0,0}
\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
  filecolor=linkcolor,urlcolor=linkcolor}
\usepackage{tikz}
\usetikzlibrary{arrows,decorations.pathmorphing}

\shorttitle{Chirping Jets from Black Hole Binaries}
\shortauthors{Kulkarni et al.}

\begin{document}

\title{Radio Crickets: Chirping Jets from Black Hole Binaries Entering
  their Gravitational Wave Inspiral} \author{Girish
  Kulkarni\altaffilmark{1} and Abraham Loeb\altaffilmark{2}}
\altaffiltext{1}{Institute of Astronomy and Kavli Institute of
  Cosmology, University of Cambridge, Madingley Road, Cambridge CB3
  0HA, UK; \email{kulkarni@ast.cam.ac.uk}} \altaffiltext{2}{Institute
  for Theory \& Computation, Harvard University, 60 Garden Street,
  Cambridge, MA 02138, USA; \email{aloeb@cfa.harvard.edu}}

\begin{abstract}
Apart from a stochastic gravitational wave (GW) background, eLISA will
also be able to detect GW emission from binary black holes.  Combined
GW and electromagnetic observations of compact binary mergers should
enable detailed studies of astrophysical processes in the strong-field
gravity regime.  Here we show that if one of component BHs of the
binary has an accretion disk with an associated jet, the morphology of
the jet can act as an electromagnetic counterpart to the GW signal
from the binary.  For binaries with mass $10^7$--$10^{10}$ M$_\odot$
at redshifts $z<0.5$, the jet precesses along a biconical surface due
to the binary's orbital motion.  When the binary enters the GW phase
of its evolution, the opening angle widens, the jet exhibits
milliarcsecond scale wiggles and the conical surface of jet precession
is twisted due to the superluminal motion.  These morphological
features can be used to constrain the dyanamical parameters of the
binary.  Monitoring of jets with these features in current and
archival data will put a lower limit on the sources that could be
detected by GW experiments such as eLISA and Pulsar Timing Arrays
(PTAs).  In the future, microarcsecond interferometry with experiments
such as SKA will increase the potential usefulness of this technique.  
\end{abstract}

\keywords{BHs}

\section{Introduction}
\label{sec:intro}

Binaries of supermassive black holes (BHs) arise naturally as a result
of mergers of galaxies in the context of the heirarchical galaxy
formation \citep{2000MNRAS.311..576K, 2002MNRAS.336L..61H,
  2008ApJ...676...33D, 2012MNRAS.422.1306K}.  After a galaxy merger,
infalling black holes lose their angular momentum in three stages
before coalescence \citep{1980Natur.287..307B, 2005LRR.....8....8M,
  2011ASL.....4..181C}.  In the first stage, the black holes sink to
the center of the gravitational potential of the merger remnant due to
dynamical friction and form a gravitationally bound binary.  The newly
formed binary then continues to lose its angular momentum and shrink
by scattering ambient gas and stars.  In the final stage of the
binary's evolution, emission of gravitational waves (GWs) becomes the
predominant mode angular momentum loss, which results in the
coalescence of the black holes.

% Any more references to add here?
General relativisitic (GR) numerical simulations can predict the
binary's evolution in its final stage up to coalescence
\citep{2005PhRvL..95l1101P, 2006PhRvL..96k1102B, 2006PhRvL..96k1101C}.
This has stimulated interest in predicting related electromagnetic
signals of black hole coalescence.  Effects considered in the
literature include afterglows due to infall of gas on the remnant
black hole \citep{2005ApJ...622L..93M}, electromagnetic variability in
the circumbinary disk due to shocks induced by the sudden mass loss of
the binary due to GW emission at coalescence
\citep{2007APS..APR.S1010B}, flares from shocked remnants of accretion
disks around recoiled remnant black hole \citep{2007PhRvL..99d1103L,
  2008ApJ...682..758S, 2009CQGra..26i4032H}, quasi-periodic
variability in gas accretion in the early stages of black hole merger
\citep{2008ApJ...672...83M, 2009CQGra..26i4032H, 2012MNRAS.427.2680K},
evolution in the profiles of broad emission lines shortly before
merger \citep{2013MNRAS.432.1468M}, characteristic evolution in
thermal emission due to circumbinary accretion
\citep{2015MNRAS.446L..36F}, brightening of accretion disk due to
viscous dissipation of gravitational waves
\citep{2008PhRvL.101d1101K}, and prompt tidal disruption of stars by
the recoiled remnant \citep{2011MNRAS.412...75S}.

% Replace `LISA' with `eLISA` or `NGO'?  Mention pulsar timing arrays?
In addition to testing GR in the strong field limit, observations of
electromagnetic counterparts will help localize sources of GWs for
future detectors such as the Evolved Laser Interferometer Space
Antenna (eLISA)\footnote{\url{https://www.elisascience.org}}, which would be
capable of observing the peak GW luminosity for BH coalescence of
$\sim 10^{57}$ erg$/$s up to large cosmological distances
\citep{2003CQGra..20S..65H, 2013CQGra..30x4009S}.  Simultaneous
detection of GW and electromagnetic signals from coalescing binaries
will yield rich scientific pay-offs \citep{2003CQGra..20S..65H,
  2005ApJ...629...15H}.  For example, by a redshift measurement of
eLISA sources could constraint cosmological parameters.  Measurement
of the black hole masses and spins from the GW signal will constrain
accretion physics during coalescence.  But even before eLISA is
launched, a search of electromagnetic signatures may point to the
expected abundance of GW sources.

In this Letter we study the morphology of a jet emitted by one of the
binary members during its final stage of evolution before coalescence,
i.e., when the binary's dynamics is dominated by GW emission.  We show
that the helical chirping morphology of the jet can be used to infer
the binary parameters.

\section{Binary parameters}

We study a BH binary formed following a gas-rich merger of two
galaxies as a result of the processes described above.  We consider
black holes with masses $M_1$ and $M_2$ in a circular orbit of radius
$a$ around each other.  Let $M=M_1+M_2$.  For simplicity, we consider
an equal-mass binary ($M_1=M_2$) on a circular orbit. We will consider
unequal black holes masses and elliptical orbits below.  The orbital
speed of each black hole is given by \citep{2010PhRvD..81d7503L}
\begin{equation}
  v = \frac{1}{2}\left(\frac{GM}{a}\right)^{1/2}\!\!\!\! = 5.8\times 10^3\, \mathrm{km~s}^{-1}M_8^{1/2}a_{16}^{-1/2},
  \label{eqn:vorb}
\end{equation}
where $a_{16}\equiv (a/10^{16}$ cm$)$ and $M_8\equiv (M/10^8$
M$_\odot)$ are the binary separation and mass,
respectively.\footnote{Note that $10^{16} \mathrm{cm}\approx 3.4\times
  10^{-3} \mathrm{pc}$.  The Schwarzchild radius of a $10^8$~M$_\odot$
  black hole is $9.6\times 10^{-6}$~pc, or $\sim 3\times 10^{13}$~cm.}
The orbital period is
\begin{equation}
  P = 2\pi\left(\frac{GM}{a^3}\right)^{-1/2}\!\!\!\! = 1.72\, \mathrm{yr}\, a_{16}^{3/2} M_8^{-1/2}.
\end{equation}
% Mention typical size of accretion disks and typical ISCO radius
% above?

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \draw (0,0) ellipse (3 and 0.5);
      \fill[black] (0,0) circle (1pt);
      \draw [-angle 45,thick] (2,-0.39) -- (3.5,-0.1);
      \draw [-angle 45,thick,dashed,dash pattern = on 4pt off 1pt] (2,-0.39) -- (2,1.5);
      \draw [-angle 45,thick] (2,-0.39) -- (2.5,0.8);
      \draw [-angle 45,thick,red] (2,-0.39) -- (3.5,0.8);
      \fill[black] (-2,0.39) circle (2pt);
      \fill[black] (2,-0.39) circle (2pt);
      \draw (4.1,-0.1) node {\large ${\mathbf{v}}_\mathrm{orb}$};
      \draw (2.7,1.0) node {\large ${\mathbf{v}}_\mathrm{jet}$};
      \draw (4.0,1.0) node {\large $\tilde{\mathbf{v}}_\mathrm{jet}$};
      \draw (2.0,1.8) node {\large $\mathbf{L}_\mathrm{orb}$};
    \end{tikzpicture}
  \end{center}
  \caption{Jet and orbital configuration. The orbital velocity
    $\mathbf{v}_\mathrm{orb}$ causes the jet to precess as the BH
    moves along the orbit. The net jet velocity is
    $\tilde{\mathbf{v}}_\mathrm{jet}={\mathbf{v}}_\mathrm{orb}+{\mathbf{v}}_\mathrm{jet}$,
    where $\mathbf{v}_\mathrm{jet}$ is its ejection velocity at the
    source.}
  \label{fig:BH_orbit}
\end{figure}

In a gas-rich galaxy merger, the BH binary loses its angular
momentum by torquing the surrounding disk through spiral arms and
expelling gas out of a region twice as large as the binary orbit
\citep{2005ApJ...622L..93M, 2007PASJ...59..427H, 2008ApJ...672...83M,
  2009MNRAS.393.1423C}.  Fresh gas re-enters the region of the binary
as a result of turbulent transport of angular momentum in the
surrounding disk \citep{2008ApJ...672...83M, 2012ApJ...755...51N,
  2012MNRAS.427.2680K, 2012A&A...545A.127R}.  Since the expelled gas
carries a specific angular momentum of $\sim va$, the coalescence time
of the binary is inversely proportional to the supply rate of fresh
gas into the binary region.  In a steady state, this supply rate,
$\dot M$ is proportional to the accretion rate of the circumbinary gas
disk.  But only a fraction of gas that enters the central gap accretes
onto the black hole and fuels quasar activity.  Therefore, it is
appropriate to express $\dot M$ in Eddington units, $\dot{\mathcal
  M}\equiv\dot M /\dot M_E$, where $\dot M_E$ is the accretion rate
required to power the limiting Eddington luminosity with a radiative
efficiency of 10\%, $\dot M_E=2.3 M_\odot\, \mathrm{yr}^{-1}M_8$.  We
therefore have \citep{2010PhRvD..81d7503L}
\begin{equation}
  \frac{a}{\dot a} = t_\mathrm{gas}\approx\frac{J}{\dot Mva}=1.1\times 10^7\,
  \mathrm{yr}\,\dot{\mathcal M}^{-1},
\end{equation}
where $J=\mu va$ is the binary's angular momentum and $\mu=M/2$ is the
binary's reduced mass.  This time scale is greater by two orders of
magnitude than the time scale for coalescence due to angular momentum
loss by GW emission, which for circular orbits is given by
\citep{1964PhRv..136.1224P}
\begin{equation}
\frac{a}{\dot a} = t_\mathrm{GW} = \frac{5}{256}\frac{c^5a^4}{G^3M^2\mu}=2.53\times 10^5\,\mathrm{yr}\,\frac{a_{16}^4}{M_8^3}.
\end{equation}
By setting $t_\mathrm{gas}=t_\mathrm{GW}$, we can get the orbital
parameters for the binary when GWs take over.  The orbital speed at
this stage of the binary's evolution is given by
\begin{equation}
  v = 3.6\times 10^3\, \mathrm{km~s}^{-1}(\dot{\mathcal{M}}M_8)^{1/8},
  \label{eqn:v}
\end{equation}
The separation between the two black holes at this instant is given by
\begin{equation}
  a = 8.3\times 10^{-3}\, \mathrm{pc}\, M_8^{3/4}\dot{\mathcal{M}}^{-1/4}.
  \label{eqn:a}
\end{equation}
Note that the separation is dependent on the gas accretion rate.  This
evolution of the binary separation and orbital velocity is shown in
Figure~\ref{fig:evolve}.

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \draw [-angle 45,dashed,thick,dash pattern = on 7pt off 2pt] (0,0) -- (3.0,0);
      \draw [-angle 45,dashed,thick,dash pattern = on 7pt off 2pt] (0,0) -- (0,4.0);
      \draw [-angle 45,thick] (0,0) -- (2.0,4.0);
      \draw [-angle 45,thick,red] (0,0) -- (5.0,4.0);
      \draw (1.0,2.0) arc (63:90:2.2);
      \draw (2.0,1.6) arc (38.6:63:2.56);
      \draw (0.5,2.5) node {\large $\chi$};
      \draw (1.90,2.3) node {\large $\psi$};
      \draw [dashed,thick,dash pattern = on 7pt off 2pt] (0,0) -- (3.0,6.0);
      \draw (0.0,4.5) node {\large $\mathbf{L}_\mathrm{orb}$};
      \draw (3.5,0.0) node {\large $\mathbf{v}_\mathrm{orb}$};
      \draw (1.0,3.0) node {\large $\mathbf{v}_\mathrm{jet}$};
      \draw (5.3,4.2) node {\large \textcolor{red}{$\mathbf{\tilde v}_\mathrm{jet}$}};
    \end{tikzpicture}
  \end{center}
  \caption{The orbital velocity $\mathbf{v}_\mathrm{orbital}$
    introduces the jet precision angle $\psi$.  Without the orbital
    motion, $\psi=0$ and the jet traces a cylindrical surface.  With
    the orbital motion, the net red velocity vector oscillates around
    the solid black vector and produces a conical surface with $\psi$
    as its half-opening angle.  Note that $v_\mathrm{orb}$ is
    exagerated here.  In practice, $v_\mathrm{orb}\ll
    v_\mathrm{jet}\sim c$.  This figure then shows that
    $\sin{\psi}\sim\tan{\psi}=v_\mathrm{orb}\cos{\chi}/v_\mathrm{jet}$.}
  \label{fig:jet}
\end{figure}

In the GW dominated phase of binary's evolution, recent simulations
show that the BHs can continue to accrete gas from the circumbinary
disk all the way until merger \citep{2015MNRAS.446L..36F}.  At this
stage, we consider a ballistic jet emanating from one of binary's
member BHs at speed $\beta=v_\mathrm{jet}/c$ at an angle $\chi$ to the
direction of the orbital angular momentum.  This geometry is depicted
in Figures~\ref{fig:BH_orbit} and \ref{fig:jet}.  As a result of the
binary orbit, the jet will wiggle on a conical surface.  (In the limit
of a small binary velocity, i.e., at large relative separations, the
jet material will move on a cylindrical surface.)  The half-opening
angle of the cone is given by \citep{1993ApJ...409..130R}
\begin{equation}
  \sin\psi = \frac{v_\mathrm{orb}\cos\chi}{v_\mathrm{jet}}.
\end{equation}
When the binary separation is where gravitational wave-induced
inspiral begins, it has a velocity given by equation~(\ref{eqn:v}),
and the separation between the two black holes at this instant is
given by equation~(\ref{eqn:a}).  Also, from equation
(\ref{eqn:vorb}), the orbital velocity of each BH evolves as
\begin{equation}
  \frac{\dot v_\mathrm{orb}}{v_\mathrm{orb}} = -\frac{1}{2}\frac{\dot a}{a}.
\end{equation}
which gives the evolution of the opening angle of the cone as 
\begin{equation}
  \dot\psi^\mathrm{obs}=\dot\psi = -\frac{1}{2\cot\psi}\frac{\dot a}{a}.
\end{equation}
This can be solved analytically to get
\begin{equation}
  \sin\psi = \sin\psi_0\left(\frac{a_0}{a}\right),
\end{equation}
where $\psi_0$ and $a_0$ are the initial values. Thus, as the binary
continues to shrink the BH orbital velocity grows and the opening
angle of the conical jet increases (cone opens wider).  The rate at
which this happens reflects the binary's dynamical evolution.
Figure~\ref{fig:evolve} shows the evolution of the half-opening angle
in the gas and GW-dominated phase.

\begin{figure}
  \begin{center}
    \includegraphics*[width=\columnwidth]{evolve.pdf}
    % gw.py (case 5) 
  \end{center}
  \caption{Evolution of the binary parameters.  As the binary
    separation $a$ decreases, the orbital velocity $v_\mathrm{orb}$
    and the opening angle $\psi$ of the precessing jet, increase.}
  \label{fig:evolve}
\end{figure}

\section{Jet precession}

We now ask if it would be possible to observe the opening of the jet
morphology.  In order to do this, it is necessary to map the above
calculations to the observer's frame \citep{1982ApJ...262..478G}.  At
a large distance compared to the binary's separation, the jet geometry
is illustrated in Figure \ref{fig:jetgeom}.  Here the source frame of
reference is shown by $x^\prime$, $y^\prime$, $z^\prime$ coordinate
systems.  The $x$, $y$, $z$ coordinate system represents the
observer's frame such that the $yz$ is the plane of the sky.  Further,
the jet is oriented such that that the $y^\prime$ axis coincides with
the $y$ axis, the original direction of jet emission is at an angle
$i$ relative to the line of sight ($x$-axis).  For simplicity, we also
assume that the $z^\prime$ axis is in the $xz$ plane.  Then the jet
will precess on a cone with a half-opening angle $\psi$ and an angular
velocity of precession $\Omega$.  In the source frame of reference,
the equation of the jet is then given by \citep{1981ApJ...246L.141H}
\begin{eqnarray}
  &v_x=&s_\mathrm{jet}\beta c\left[\sin\psi\sin i\cos\Omega t+\cos\psi\cos i\right],\\
  &v_y=&s_\mathrm{jet}\beta c\sin\psi\sin\Omega t,~\,\mathrm{and}\\
  &v_z=&s_\mathrm{jet}\beta c\left[\cos\psi\sin i-\sin\psi\cos i\cos\Omega t\right]. 
\end{eqnarray}
Here $s_\mathrm{jet}=\pm 1$ for forward and backward jets,
respectively.  The coordinates of a jet particle in the sky is then
given by $z=v_zt$ and $y=v_yt$.  However, because of superluminal
motion, the observer will see this on the sky as
\begin{equation}
  z=v_zt/(1-v_x/c)
\end{equation}
and
\begin{equation}
  y=v_yt/(1-v_x/c)
\end{equation}
respectively.  In angular coordinates, the motion on the sky is given
by $\phi_z=z/d$ and $\phi_y=y/d$ where $d$ is the angular diameter
distance of the binary.  Because of superluminal motion, the forward
jet is stretched while the backward jet is compressed.

\begin{figure}
  \begin{center}
    \begin{tikzpicture}
      \draw [thick,color=brown] (0,0) -- (-4.0,0);
      \draw [thick,color=brown] (0,0) -- (0,4.0);
      \draw [thick,color=brown] (0,0) -- (2.6,-3.0);

      \draw [dashed,thick,dash pattern = on 7pt off 2pt,color=red] (0.05,0.0) -- (0.05,4.0);
      \draw [dashed,thick,dash pattern = on 7pt off 2pt,color=red] (0,0.0) -- (3.46,-2.0);
      \draw [dashed,thick,dash pattern = on 7pt off 2pt,color=red] (0,0.0) -- (-3.87,-1.0);

      \draw [thick,rotate around={-75:(-2.95,-0.76)}] (-2.95,-0.76) ellipse (0.55 and 0.2);
      \draw [thick,rotate around={12:(3.04,0.24)}] (3.04,0.238) arc (-90:90:0.2 and 0.55);
      \draw [thick,color=gray,rotate around={12:(3.04,0.24)},dashed](3.04,0.238) arc (270:90:0.2 and 0.55);

      \draw [thick] (-2.82,-1.3) -- (2.82,1.3);
      \draw [thick] (-3.09,-0.24) -- (3.09,0.24);

      \draw (0.1,4.3) node {\large \textcolor{brown}{$y$},\textcolor{red}{$\,y^\prime$}};
      \draw (-4.2,0) node {\textcolor{brown}{\large $z$}};
      \draw (2.75,-3.17) node {\textcolor{brown}{\large $x$}};
      \draw (3.1,-3.57) node {\large [line of sight]};
      \draw (-4.1,-1.059) node {\large \textcolor{red}{$z^\prime$}};
      \draw (3.7,-2.1) node {\textcolor{red}{\large $x^\prime$}};

      \draw [thick,color=brown] (-1.0,-0.258) arc (230:283:1.8);
      \draw (-0.5,-1.0) node {\textcolor{brown}{\large $i$}};

      \draw (-2.0,-0.92) arc (204.75:194.48:2.2015);
      \draw (-2.0,-1.3) node {\large $\psi$};

    \end{tikzpicture}
  \end{center}
  \caption{Geometry of the jet.  The observer's sky is in the $yz$
    plane, so that the line of sight is along the $x$-axis.  The jet
    precesses around the $z^\prime$ axis with half-opening angle
    $\psi$. The $y$ and $y^\prime$ axes overlap, and $z$, $x$, and
    $z^\prime$ axes are co-planer.  The angle between the $z^\prime$
    axis and the line of sight is $i$.  This geometry is valid in the
    limit $a\ll z^\prime\psi$.}
  \label{fig:jetgeom}
\end{figure}

Finally, we assume that the emission from each jet particle is
optically thin at the frequency of observation with a spectrum in the
rest frame of the particle given by $P(\nu)\propto \nu^{-\alpha}$
\citep{1982ApJ...262..478G}.  The observed flux density $S(\nu)$ in erg
cm$^{-2}$ Hz$^{-1}$ is given by
\begin{equation}
  S(\nu)=S_r(\nu)D^{3+\alpha},
\end{equation}
where $D$ is the Doppler shift parameter
$[\gamma(1-\beta\cos\phi)]^{-1}$ and $S_r$ is the rest frame flux
density.  Further to consider the effects of jet evolution we assume
that $S_r$ decreases with time according to a simple power law
\begin{equation}
  S_r\propto t^{-\delta}.
\end{equation}
We assume $\delta=1$ and $\alpha=1$, hereafter. 
% Need to justify parameter values.
% Comment on relativistic corrections

\begin{figure*}
\begin{center}
  \begin{tabular}{cc}
    \includegraphics*[scale=0.45]{{jet_i40_beta0.90_mdot1.00_full}.pdf} &
    \includegraphics*[scale=0.45]{jet_check_zoom1.pdf}
  \end{tabular}
\end{center}
\caption{The left panel shows forward and backward jets for an
  equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  We have
  assumed $i=40^\circ$, $\beta=0.9$, and $d = 100$~Mpc.  The jets are
  twisted because of the apparant superluminal motion.  Also evident
  is the opening of the jet and the radio jet chirp.  Right panel
  shows normalised brightness of the inner 30~mas region of the
  forward jet.}
\label{fig:jet_mdot1.0}
\end{figure*}

\section{Flagging GW binaries and probing GR}

As the binary evolves, its increasing orbital period and speed, and
the apparant superluminal motion effect, together result in a jet
morphology with an increasing opening angle and increasingly rapid
periodic variation together with a twisted shape.
Figure~\ref{fig:jet_mdot1.0} shows these three features for an
equal-mass binary with total mass $M=10^{10}$~M$_\odot$.  For
concreteness, we have assumed $i=40^\circ$ and $\beta=0.9$.  The
angular diameter distance to the binary is assumed to be $d =
100$~Mpc, which corresponds to redshift $z\sim 0.3$.  The forward jet
streched on the sky while the backward jet is compressed due to the
superluminal motion.  The right panel of Figure~\ref{fig:jet_mdot1.0}
zooms into the inner region of the forward jet and shows that the
binary creates wiggles in the jet morphology of the order of a
milli-arcsecond.  The duration of the binary evolution captured by
Figure~\ref{fig:jet_mdot1.0} is 100 yr.

% Add discussion about the chirp of the signal

\begin{table}
\begin{center}
\begin{tabular}{c l}
\hline 
Parameter & Description \\ 
\hline 
$M$ & Total mass \\
$\zeta$ & Mass asymmetry \\
$\chi$ & Intrinsic jet direction \\
$e$ & Orbital eccentricity \\
$\alpha$ & SED slope \\
$\delta$ & Time evolution of brightness evolution \\
$\beta$ & Jet speed \\
$i$ & Orbital inclination \\ 
$\dot{\mathcal M}$ & Gas accretion rate\\
\hline
\end{tabular}
\end{center}
\caption{Parameters of the problem.}
\label{table:params}
\end{table}

\section{Conclusion}

Using a model of jet precession in a binary BH, we have shown that the
jet morphology of a binary that is entering its GW-induced phase of
inspiral shows a distinct chirping morphology. Jets in these binaries
have a biconical morphology.  After the binary enters the GW phase,
the opening angle of cone rapidly increases within a distance of tens
of milli-arcseconds.  In addition, the periodicity of the jet's motion
on the conical surface increases as the orbital period of the binary
goes up in this stage of its evolution.  Finally, due to apparent
superluminal motion and the widening opening angle, the jet is twisted
into a peculiar morphology.  The jet geometry can be used to infer the
set of dynamical and geometrical parameters, listed in
Table~\ref{table:params}.

% Give redshift numbers
These milli-arcsecond features can be resolved using long baseline
radio interferometry \citep{2001ChJAA...1..236Q}.  At higher
redshifts, jets are fainter and the wiggles become more difficult to
resolve.  However, it might be possible to resolve binaries with very
unequal mass ratios, eccentric orbits, or suitable inclinations.  The
future Square Kilometer Array (SKA) should make microarcsecond
resolution possible, which could increase the detection rate of these
objects \citep{2014arXiv1412.5971P}.

The jet morphology acts as an electromagnetic counterpart for chirping
gravitational wave sources.  BH binaries are expected to be the
brightest GW sources in the universe.  Current ground based GW
detectors such as LIGO are sensitive to low-mass BH binaries at high
redshift, which would be difficult to detect with this method.
However, future detectors such as eLISA are sensitvie to GW emission
from $10^7$--$10^9$ M$_\odot$ BHs \citep{2013CQGra..30x4009S}.  Pulsar
Timing Arrays are sensitive to even higher binary masses.  Monitoring
of jet morphologies with features outlined in this paper will put a
lower bound on the expected abundance of bright GW sources for this
experiments. Moreover, combined GW and electromagnetic observations of
compact binary mergers should enable detailed studies of astrophysical
processes in the strong-field gravity regime
\citep{2005ApJ...629...15H}.

% Mention old observations which show similar features
% There is a todo in a PS email from Avi

\section*{Acknowledgments}

We acknowledge useful discussions with Enrico Ramirez-Ruiz and Alberto
Rorai.  This work was supported in part by NSF grant AST-1312034 (for
A.~L.).

\bibliography{refs}
\end{document}